---
title: "CO2-Emission Modeling"
author: "Andrew Mashhadi"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the Data & Libraries

```{r, echo = FALSE, warning = FALSE, error = FALSE}
library(DAAG)
library(dplyr)
library(effects)
library(car)
library(corrplot) 
library(psych)
library(caret)
library(readr)
library(ggplot2)
```


# Clean the data as recommended by the EDA, using One-Hot-Encoding

```{r}

co2_df <- na.omit(read.csv("dataset/CO2_emissions.csv"))

# group the manufacturers into less categories (with commonalities)
co2_df <- co2_df %>% 
          mutate(econ.class = case_when(
          Make == "FORD" | Make == "RAM" | 
          Make == "DODGE" | Make == "CHRYSLER" |
          Make == "CHEVROLET" | Make == "GMC" |
          Make == "JEEP" |
          Make == "HONDA" | Make == "HYUNDAI" | 
          Make == "KIA" | Make == "MITSUBISHI" |
          Make == "NISSAN" | Make == "SCION" |  
          Make == "SUBARU" | Make == "TOYOTA" |
          Make == "MAZDA" |
          Make == "VOLKSWAGON" | Make == "VOLVO" | 
          Make == "SMART" | Make == "FIAT" ~ "Economy",
          Make == "BUICK" | Make == "LINCOLN" | 
          Make == "CADILLAC" | Make == "SRT" |
          Make == "JAGUAR" |
          Make == "ACURA" | Make == "GENESIS" | 
          Make == "INFINITI" | Make == "LEXUS" |
          Make == "AUDI" | Make == "ALFA ROMERO" |
          Make == "BMW" | Make == "MINI" |
          Make == "MERCEDES-BENZ" | Make == "LAND ROVER" ~ "Premium",
          Make == "ASTON MARTIN" | Make == "BENTLEY" | 
          Make == "BUGATTI" | Make == "LAMBORCHINI" | 
          Make == "PORCHSE" | Make == "ROLLS-ROYCE" ~ "Luxury")) %>%
          select(-c(Make))


# simplified categories
co2_df <- co2_df %>% 
          mutate(drive = case_when(
          grepl("4WD", Model, ignore.case = TRUE) |
          grepl("4X4", Model, ignore.case = TRUE) |
          grepl("AWD", Model, ignore.case = TRUE) |
          grepl("xDrive", Model, ignore.case = TRUE) ~ "All-Wheel Drive",
          TRUE ~ "Two-Wheel Drive")) %>%
          select(-c(Model))


co2_df <- co2_df %>% 
          mutate(Vehicle.Class = case_when(
          Vehicle.Class == "COMPACT" | Vehicle.Class == "SUBCOMPACT" |
          Vehicle.Class == "MINICOMPACT" | Vehicle.Class == "TWO-SEATER" ~ "COMPACT",
          Vehicle.Class == "MID-SIZE" | Vehicle.Class == "STATION WAGON - MID-SIZE" | 
          Vehicle.Class == "STATION WAGON - SMALL"~ "MID-SIZE",
          Vehicle.Class == "FULL-SIZE" ~ "FULL-SIZE",
          Vehicle.Class == "MINIVAN" | Vehicle.Class == "VAN - CARGO" |
          Vehicle.Class == "VAN - PASSENGER"  |
          Vehicle.Class == "PICKUP TRUCK - SMALL" | 
          Vehicle.Class == "PICKUP TRUCK - STANDARD" |
          Vehicle.Class == "SUV - SMALL" | 
          Vehicle.Class == "SUV - STANDARD" ~ "LARGE VEHICLE"))


# I break up the transmission categories by transmission type and number
# of gears
co2_df$transmission.type <- factor(sub("[0-9^]+", "", co2_df$Transmission))


co2_df <- co2_df %>% 
          mutate(transmission.type = case_when(
          transmission.type == "M" ~ "Man",
          TRUE ~ "Auto"))

co2_df$gear.num <- as.numeric(gsub("\\D+", "", co2_df$Transmission))
co2_df$gear.num[is.na(co2_df$gear.num)] = 0 # continuous technically has 0 gears


co2_df <- co2_df %>% 
          select(-c(Transmission))

# remove the one observation with natural gas (small sample size not good for inference)
co2_df <- co2_df[co2_df$Fuel.Type != "N", ]
co2_df <- co2_df %>% 
          mutate(Fuel.Type = case_when(
          Fuel.Type == "X" ~ "Regular",
          TRUE ~ "Non-Regular"))


# remove the mixed fuel consumption (PCA and FA should find a linear combination anyway) 
co2_df <- co2_df %>% 
          select(-c(Fuel.Consumption.Comb, Fuel.Consumption.Comb.MPG))

co2_df <- na.omit(co2_df)


# convert types to factors
co2_df$econ.class <- factor(co2_df$econ.class)
co2_df$drive <- factor(co2_df$drive)
co2_df$Vehicle.Class <- factor(co2_df$Vehicle.Class)
co2_df$Fuel.Type <- factor(co2_df$Fuel.Type)
co2_df$transmission.type <- factor(co2_df$transmission.type)

# make dummy variables for PCA and FA
dummy <- dummyVars(" ~ .", data=co2_df, fullRank = TRUE)
co2_df <- data.frame(predict(dummy, newdata = co2_df))



```



# Train/Test Split

```{r}

# to reproduce
set.seed(5)

#use 80% of dataset as training set and 20% as test set
sample <- sample(c(TRUE, FALSE), nrow(co2_df), replace=TRUE, prob=c(0.8,0.2))
train  <- co2_df[sample, ]
test   <- co2_df[!sample, ]


# 
normParam <- preProcess(train[, -c(1:3, 6, 9, 10:13)])
train[, -c(1:3, 6, 9, 10:13)] <- predict(normParam, train[, -c(1:3, 6, 9, 10:13)])
test[, -c(1:3, 6, 9, 10:13)] <- predict(normParam, test[, -c(1:3, 6, 9, 10:13)])



```


# PCA + Multiple Linear Regression

## PCA

```{r}

# remove response variable before calculating the sample principal components (CO2 Emission) 
PCs <- prcomp(subset(train, select = -c(CO2.Emissions)), center=FALSE, scale = FALSE)

# output summary of the sample principal components
summary(PCs)

# print the eigenvector (loadings)
PCs$rotation[, 1:6]

# scree plot
var_explained = PCs$sdev^2 / sum(PCs$sdev^2)

plot(1:length(PCs$sdev), var_explained,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot",
     xlab = "Sample Principal Component Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width


```


## Linear Regression Model with PCs


```{r}

train.control <- trainControl(method = "cv", number = 5)

train.PCs <- data.frame(PCs$x, CO2.Emissions = train$CO2.Emissions)

# examine the 10-fold CV performance for different number of PCs
cv.results <- data.frame()
for (k in 1:(ncol(train.PCs)-1)){
  
  cv.pca.lm <- train(CO2.Emissions ~ ., 
                      data = train.PCs[, c(1:k, 14)], 
                      method = "lm", 
                      trControl = train.control)
  # save cv results
  cv.results <- rbind(cv.results, list(k=k, RMSE=cv.pca.lm$results$RMSE, 
                                       Rsq=cv.pca.lm$results$Rsquared, 
                                       MAE=cv.pca.lm$results$MAE))
  print(cv.pca.lm)
}

# display results
cv.results

par(mfrow=c(1, 2))
plot(RMSE ~ k, data=cv.results, 
     main="5-Fold CV Results",
     xlab="# Principal Components", 
     col= "blue", 
     pch = 19)
grid(nx = NULL, ny = NULL,
     lty = 1,      # Grid line type
     col = "gray", # Grid line color
     lwd = 1)      # Grid line width
plot(Rsq ~ k, data=cv.results, 
     main="5-Fold CV Results",
     xlab="# Principal Components", 
     col= "green", 
     pch = 19)
grid(nx = NULL, ny = NULL,
     lty = 1,      # Grid line type
     col = "gray", # Grid line color
     lwd = 1)      # Grid line width

```

```{r}
k.opt <- 6

# re-train full model
pca.lm <- lm(CO2.Emissions ~ ., data = train.PCs[, c(1:k.opt, 14)])
summary(pca.lm)

par(mfrow=c(2,2))
plot(pca.lm, which = c(1, 2, 4, 5))
```



```{r}

# Setup test set using training loadings
x <- as.matrix(subset(test, select = -c(CO2.Emissions)))
test.PCs <- x %*% PCs$rotation[, 1:k.opt]
test.PCs <- data.frame(test.PCs, CO2.Emissions = test$CO2.Emissions)

preds <- predict(pca.lm, newdata = test.PCs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))

# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)

```


# FA + Multiple Linear Regression

## Factor Analysis


```{r}

# remove the response variable for FA
preds <- subset(train, select=-c(CO2.Emissions))

R <- cor(preds)


# principal component solution
fa_pc <- principal(R, nfactors = 12, rotate = "varimax")
fa_pc$loadings



# maximum likelihood solution
fa_mle <- fa(R, nfactors = 12, fm = "mle", rotate = "varimax")
fa_mle$loadings


# scree plots
var_exp_pc <- c()
for (i in 1:12) {
    var_exp_pc <- c(var_exp_pc, sum((fa_pc$loadings[, i])^2) / sum(diag(R)))
}


var_exp_mle <- c()
for (i in 1:12) {
    var_exp_mle <- c(var_exp_mle, sum((fa_mle$loadings[, i])^2) / sum(diag(R)))
}


par(mfrow = c(1, 2))

plot(1:12, var_exp_pc,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot for PC Method",
     xlab = "Factor Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width



plot(1:12, var_exp_mle,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot for MLE Method",
     xlab = "Factor Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width



```


## Linear Regression with the Factor Scores


```{r}
Pz <- diag(fa_pc$uniquenesses)
Lz <- fa_pc$loadings[, 1:11]
z <- train[, -9]


# wls factor score estimations
f_uls <- solve(t(Lz) %*% Lz) %*% t(Lz) %*% t(z) 

head(t(f_uls), n = 5)

```

```{r}

train.Fs <- data.frame(t(f_uls), CO2.Emissions = train$CO2.Emissions)

# re-train full model
fa.pc.lm <- lm(CO2.Emissions ~ ., data = train.Fs)
summary(fa.pc.lm)

par(mfrow=c(2,2))
plot(fa.pc.lm, which = c(1, 2, 4, 5))

```


```{r}

Pz <- diag(fa_pc$uniquenesses)
Lz <- fa_pc$loadings[, 1:11]
z0 <- test[, -9]


# uls factor score estimations
f_uls <- solve(t(Lz) %*% Lz) %*% t(Lz) %*% t(z0)

test.Fs <- data.frame(t(f_uls), CO2.Emissions = test$CO2.Emissions)

preds <- predict(fa.pc.lm, newdata = test.Fs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))

# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)

```



```{r}
Pz <- diag(fa_mle$uniquenesses)
Lz <- fa_mle$loadings[, 1:8]
z <- train[, -9]

# wls factor score estimations
f_wls <- solve(t(Lz) %*% solve(Pz) %*% Lz) %*% t(Lz) %*% solve(Pz) %*% t(z) 

head(t(f_wls), n = 5)

```


```{r}

train.Fs <- data.frame(t(f_wls), CO2.Emissions = train$CO2.Emissions)

# re-train full model
fa.mle.lm <- lm(CO2.Emissions ~ ., data = train.Fs)
summary(fa.mle.lm)

par(mfrow=c(2,2))
plot(fa.mle.lm, which = c(1, 2, 4, 5))

```

```{r}

Pz <- diag(fa_mle$uniquenesses)
Lz <- fa_mle$loadings[, 1:8]
z0 <- test[, -9]

# wls factor score estimations
f_wls <- solve(t(Lz) %*% solve(Pz) %*% Lz) %*% t(Lz) %*% solve(Pz) %*% t(z0) 

test.Fs <- data.frame(t(f_wls), CO2.Emissions = test$CO2.Emissions)

preds <- predict(fa.mle.lm, newdata = test.Fs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))

# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)

```




# Alternative Correlation Structure Approaches


```{r}


co2_df <- na.omit(read.csv("dataset/CO2_emissions.csv"))

# group the manufacturers into less categories (with commonalities)
co2_df <- co2_df %>% 
          mutate(econ.class = case_when(
          Make == "FORD" | Make == "RAM" | 
          Make == "DODGE" | Make == "CHRYSLER" |
          Make == "CHEVROLET" | Make == "GMC" |
          Make == "JEEP" |
          Make == "HONDA" | Make == "HYUNDAI" | 
          Make == "KIA" | Make == "MITSUBISHI" |
          Make == "NISSAN" | Make == "SCION" |  
          Make == "SUBARU" | Make == "TOYOTA" |
          Make == "MAZDA" |
          Make == "VOLKSWAGON" | Make == "VOLVO" | 
          Make == "SMART" | Make == "FIAT" ~ "Economy",
          Make == "BUICK" | Make == "LINCOLN" | 
          Make == "CADILLAC" | Make == "SRT" |
          Make == "JAGUAR" |
          Make == "ACURA" | Make == "GENESIS" | 
          Make == "INFINITI" | Make == "LEXUS" |
          Make == "AUDI" | Make == "ALFA ROMERO" |
          Make == "BMW" | Make == "MINI" |
          Make == "MERCEDES-BENZ" | Make == "LAND ROVER" ~ "Premium",
          Make == "ASTON MARTIN" | Make == "BENTLEY" | 
          Make == "BUGATTI" | Make == "LAMBORCHINI" | 
          Make == "PORCHSE" | Make == "ROLLS-ROYCE" ~ "Luxury")) %>%
          select(-c(Make))


# simplified categories
co2_df <- co2_df %>% 
          mutate(drive = case_when(
          grepl("4WD", Model, ignore.case = TRUE) |
          grepl("4X4", Model, ignore.case = TRUE) |
          grepl("AWD", Model, ignore.case = TRUE) |
          grepl("xDrive", Model, ignore.case = TRUE) ~ "All-Wheel Drive",
          TRUE ~ "Two-Wheel Drive")) %>%
          select(-c(Model))


co2_df <- co2_df %>% 
          mutate(Vehicle.Class = case_when(
          Vehicle.Class == "COMPACT" | Vehicle.Class == "SUBCOMPACT" |
          Vehicle.Class == "MINICOMPACT" | Vehicle.Class == "TWO-SEATER" ~ "COMPACT",
          Vehicle.Class == "MID-SIZE" | Vehicle.Class == "STATION WAGON - MID-SIZE" | 
          Vehicle.Class == "STATION WAGON - SMALL"~ "MID-SIZE",
          Vehicle.Class == "FULL-SIZE" ~ "FULL-SIZE",
          Vehicle.Class == "MINIVAN" | Vehicle.Class == "VAN - CARGO" |
          Vehicle.Class == "VAN - PASSENGER"  |
          Vehicle.Class == "PICKUP TRUCK - SMALL" | 
          Vehicle.Class == "PICKUP TRUCK - STANDARD" |
          Vehicle.Class == "SUV - SMALL" | 
          Vehicle.Class == "SUV - STANDARD" ~ "LARGE VEHICLE"))


# I break up the transmission categories by transmission type and number
# of gears
co2_df$transmission.type <- factor(sub("[0-9^]+", "", co2_df$Transmission))


co2_df <- co2_df %>% 
          mutate(transmission.type = case_when(
          transmission.type == "M" ~ "Man",
          TRUE ~ "Auto"))

co2_df$gear.num <- as.numeric(gsub("\\D+", "", co2_df$Transmission))
co2_df$gear.num[is.na(co2_df$gear.num)] = 0 # continuous technically has 0 gears


co2_df <- co2_df %>% 
          select(-c(Transmission))

# remove the one observation with natural gas (small sample size not good for inference)
co2_df <- co2_df[co2_df$Fuel.Type != "N", ]
co2_df <- co2_df %>% 
          mutate(Fuel.Type = case_when(
          Fuel.Type == "X" ~ "Regular",
          TRUE ~ "Non-Regular"))

# remove the mixed fuel consumption (PCA and FA should find a linear combination anyway) 
co2_df <- co2_df %>% 
          select(-c(Fuel.Consumption.Comb, Fuel.Consumption.Comb.MPG))

co2_df <- na.omit(co2_df)


# convert types to factors
co2_df$econ.class <- factor(co2_df$econ.class)
co2_df$drive <- factor(co2_df$drive)
co2_df$Vehicle.Class <- factor(co2_df$Vehicle.Class)
co2_df$Fuel.Type <- factor(co2_df$Fuel.Type)
co2_df$transmission.type <- factor(co2_df$transmission.type)

summary(co2_df)


```


# Train/Test Split

```{r, warning=FALSE}

# to reproduce
set.seed(5)

#use 80% of dataset as training set and 20% as test set
sample <- sample(c(TRUE, FALSE), nrow(co2_df), replace=TRUE, prob=c(0.8,0.2))
train  <- co2_df[sample, ]
test   <- co2_df[!sample, ]

# scale only the non-factors
normParam <- preProcess(train[, -7])
train[, -7] <- predict(normParam, train[, -7])
test[, -7] <- predict(normParam, test[, -7])

# prepare train and test sets
train$econ.class <- as.integer(train$econ.class)
train$drive <- as.integer(train$drive) - 1
train$Vehicle.Class <- as.integer(train$Vehicle.Class)
train$Fuel.Type <- as.integer(train$Fuel.Type)
train$transmission.type <- as.integer(train$transmission.type) - 1

test$econ.class <- as.integer(test$econ.class)
test$drive <- as.integer(test$drive) - 1
test$Vehicle.Class <- as.integer(test$Vehicle.Class)
test$Fuel.Type <- as.integer(test$Fuel.Type)
test$transmission.type <- as.integer(test$transmission.type) - 1

# mixed correlations
R <- mixedCor(train[, -7], c = c(2, 3, 5, 6, 10), d = c(4, 8, 9), p = c(1, 7))

```


# PCA + Multiple Linear Regression

## PCA

```{r}

PCs <- princomp(covmat=R$rho)

# output summary of the sample principal components
summary(PCs)

# print the eigenvector (loadings)
PCs$loadings[, 1:6]

# scree plot
var_explained = PCs$sdev^2 / sum(PCs$sdev^2)

plot(1:length(PCs$sdev), var_explained,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot",
     xlab = "Sample Principal Component Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width


```


## Linear Regression Model with PCs

```{r}

# Setup train set using training loadings
x <- as.matrix(subset(train, select = -c(CO2.Emissions)))
train.PCs <- x %*% PCs$loadings
train.PCs <- data.frame(train.PCs, CO2.Emissions = train$CO2.Emissions)

train.control <- trainControl(method = "cv", number = 5)

# examine the 10-fold CV performance for different number of PCs
cv.results <- data.frame()
for (k in 1:(ncol(train.PCs)-1)){
  
  cv.pca.lm <- train(CO2.Emissions ~ ., 
                      data = train.PCs[, c(1:k, 11)], 
                      method = "lm", 
                      trControl = train.control)
  # save cv results
  cv.results <- rbind(cv.results, list(k=k, RMSE=cv.pca.lm$results$RMSE, 
                                       Rsq=cv.pca.lm$results$Rsquared, 
                                       MAE=cv.pca.lm$results$MAE))
  print(cv.pca.lm)
}

# display results
cv.results

par(mfrow=c(1, 2))
plot(RMSE ~ k, data=cv.results, 
     main="5-Fold CV Results",
     xlab="# Principal Components", 
     col= "blue", 
     pch = 19)
grid(nx = NULL, ny = NULL,
     lty = 1,      # Grid line type
     col = "gray", # Grid line color
     lwd = 1)      # Grid line width
plot(Rsq ~ k, data=cv.results, 
     main="5-Fold CV Results",
     xlab="# Principal Components", 
     col= "green", 
     pch = 19)
grid(nx = NULL, ny = NULL,
     lty = 1,      # Grid line type
     col = "gray", # Grid line color
     lwd = 1)      # Grid line width


```


```{r}

# re-train full model
pca.lm <- lm(CO2.Emissions ~ ., data = train.PCs[, c(1:6, 11)])
summary(pca.lm)

par(mfrow=c(2,2))
plot(pca.lm, which = c(1, 2, 4, 5))
```


```{r}

# Setup test set using training loadings
x <- as.matrix(subset(test, select = -c(CO2.Emissions)))
test.PCs <- x %*% PCs$loadings[, 1:6]
test.PCs <- data.frame(test.PCs, CO2.Emissions = test$CO2.Emissions)

preds <- predict(pca.lm, newdata = test.PCs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))


# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)
```

# FA + Multiple Linear Regression

## Factor Analysis

```{r, warning=FALSE}

# principal component solution
fa_pc <- principal(R$rho, nfactors = 10, rotate='varimax')
fa_pc$loadings

# maximum likelihood solution
fa_mle <- fa(R$rho, nfactors = 9, fm = "mle", rotate = 'varimax')
fa_mle$loadings


# scree plots
var_exp_pc <- c()
for (i in 1:10) {
    var_exp_pc <- c(var_exp_pc, sum((fa_pc$loadings[, i])^2) / sum(diag(R$rho)))
}


var_exp_mle <- c()
for (i in 1:9) {
    var_exp_mle <- c(var_exp_mle, sum((fa_mle$loadings[, i])^2) / sum(diag(R$rho)))
}


par(mfrow = c(1, 2))

plot(1:10, var_exp_pc,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot for PC Method",
     xlab = "Factor Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width



plot(1:9, var_exp_mle,
     type = "l", lty = 1, lwd = 3,
     main = "Scree Plot for MLE Method",
     xlab = "Factor Number",
     ylab = "Variance Explained")
grid(nx = NULL, ny = NULL,
     lty = 2,      # Grid line type
     col = "gray", # Grid line color
     lwd = 2)      # Grid line width



```

## Linear Regression with the Factor Scores


```{r}
Pz <- diag(fa_pc$uniquenesses)
Lz <- fa_pc$loadings[, 1:8]
z <- train[, -7]


# uls factor score estimations
f_uls <- solve(t(Lz) %*% Lz) %*% t(Lz) %*% t(z) 

head(t(f_uls), n = 5)

```

```{r}

train.Fs <- data.frame(t(f_uls), CO2.Emissions = train$CO2.Emissions)

# re-train full model
fa.pc.lm <- lm(CO2.Emissions ~ ., data = train.Fs)
summary(fa.pc.lm)

par(mfrow=c(2,2))
plot(fa.pc.lm, which = c(1, 2, 4, 5))

```

```{r}

Pz <- diag(fa_pc$uniquenesses)
Lz <- fa_pc$loadings[, 1:8]
z0 <- test[, -7]


# uls factor score estimations
f_uls <- solve(t(Lz) %*% Lz) %*% t(Lz) %*% t(z0)

test.Fs <- data.frame(t(f_uls), CO2.Emissions = test$CO2.Emissions)

preds <- predict(fa.pc.lm, newdata = test.Fs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))

# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)

```


```{r}
Pz <- diag(fa_mle$uniquenesses)
Lz <- fa_mle$loadings[, 1:8]
z <- train[, -7]

# wls factor score estimations
f_wls <- solve(t(Lz) %*% solve(Pz) %*% Lz) %*% t(Lz) %*% solve(Pz) %*% t(z) 

head(t(f_wls), n = 5)

```


```{r}

train.Fs <- data.frame(t(f_wls), CO2.Emissions = train$CO2.Emissions)

# re-train full model
fa.mle.lm <- lm(CO2.Emissions ~ ., data = train.Fs)
summary(fa.mle.lm)

par(mfrow=c(2,2))
plot(fa.mle.lm, which = c(1, 2, 4, 5))

```


```{r}

Pz <- diag(fa_mle$uniquenesses)
Lz <- fa_mle$loadings[, 1:8]
z0 <- test[, -7]

# wls factor score estimations
f_wls <- solve(t(Lz) %*% solve(Pz) %*% Lz) %*% t(Lz) %*% solve(Pz) %*% t(z0) 

test.Fs <- data.frame(t(f_wls), CO2.Emissions = test$CO2.Emissions)

preds <- predict(fa.mle.lm, newdata = test.Fs)

## RMSE
sqrt(mean((preds - test$CO2.Emissions)^2))

# R-Sq
RSS <- sum((preds - test$CO2.Emissions)^2)
TSS <- sum((test$CO2.Emissions-mean(test$CO2.Emissions))^2)
1 - (RSS/TSS)

```


